---
title: "Final_Project_FyonaSun"
author: "Fyona Sun"
date: "2/10/2020"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
auction=read_csv('./online-auctions-dataset/auction.csv')
head(auction)
dim(auction)
sapply(auction, class)
```

The data set includes 10681 observations and 9 variables
auctionid: unique identifier of an auction
bid: the proxy bid placed by a bidder
bidtime: the time in days that the bid was placed, from the start of the auction
bidder: eBay username of the bidder
bidderrate: eBay feedback rating of the bidder
openbid: the opening bid set by the seller
price: the closing price that the item sold for (equivalent to the second highest bid + an increment)
item: auction item
auction_type

Understanding eBay auction policy
It's important to understand how eBay auction policy works before conducting further analysis. I found this explanation easy to understand.

The website helps to answer the following questions:

What's the relationship between bids and bidtime? Poeple have to bid higher than current price, which is the second highest bid+increment. If someone places a bid much higher than current price, eBay will only use part of his/her bid(second highest bid+increment) as the new current price. People can bid as long as their bid is higher than current price. Therefore, as bidtime increases, we sometimes see bids that are lower than previous bids. People will be notified when they are outbid, so they may increase their max bids to remain the higher bidder.

What's the relationship between bids and closing price? Closing price=second highest bid+increment or highest bid in some cases when second highest bid+increment is greater than the thrid highest bid. For example, if third highest bid is 8 and second highest bid is 10, you can bid 10.01 and if no one's bid is higher than yours, you get the item by paying 10.01 instead of 10+increment(which is 10.5)

Person who placed the highest bid gets the item but how much he/she needs to pay in most cases depends on the second highest bid.
## Exploring the dataset

```{r}
## function bid.incr(price)
bid.incr <- 
       function(price) 
           { bidint <- c(0,0.99,4.99,24.99,99.99,249.99,499.99,999.99,2499.99,4999.99)
               ## interval of prices
             inc <- c(0.05,0.25,0.5,1,2.5,5,10,25,50,100)
               ## increments of price for particular interval
             inc[findInterval(price,bidint)]
           }
############################################################
##########            make livebids              ###########
############################################################

## function recover.livebids(bids,bidder,openbid,winningbid)
## This function reconstructs the sequence of livebids by 
## inputing the raw bids,bidder IDs and opening bid

recover.livebids <- 
       function(bids,bidder,openbid,winningbid)
           { nbid <- length(bids)
             if (nbid==1) { live <- openbid
                            live
                           }
             else {
                   if (nbid==2) {live <- numeric(nbid)
                                 live[1] <- openbid
                                 live[nbid] <- winningbid
                                 live}

                   else { live <- numeric(nbid)
                          maxbid <- bids[1]
                          live[1] <- openbid
                          live[nbid] <- winningbid
                          currentprice <- bids[1]
                          maxbidder <- bidder[1]
                          for (i in 2:(nbid-1)) 
                                { newbid <- bids[i]
                                  newbidder <- bidder[i]
                                  if (newbid > maxbid) { if (!(identical(newbidder,maxbidder))) 
                                                             { incr <- bid.incr(maxbid)
                                                               if (newbid < maxbid+incr)  currentprice <- newbid
                                                               else currentprice <- maxbid+incr
                                                               maxbid <- newbid         
                                                               maxbidder <- newbidder
                                                              } 
                                                         else {maxbid <- newbid 
                                                               maxbidder <- maxbidder
                                                               }
                                                       }
                                  else {if (newbid < maxbid) { incr <- bid.incr(newbid)
                                                               if (newbid+incr>=maxbid) currentprice <- maxbid
                                                               else currentprice <- newbid+incr
                                                               maxbid <- maxbid
                                                               maxbidder <- maxbidder
                                                              }
                                         else { if (!(identical(newbidder,maxbidder))) 
                                                                            {currentprice <- maxbid
                                                                             maxbid <- maxbid
                                                                             maxbidder <- maxbidder
                                                                            }
                                                else {currentprice <- currentprice
                                                      maxbid <- maxbid
                                                      maxbidder <- maxbidder
                                                      }                                              
                                              }
                                        }
                                 live[i] <- currentprice
                               }
                         live
                        }
                  }
             }

```

```{r}
auction=read_csv('./online-auctions-dataset/auction.csv')
swarovski=read_csv('./online-auctions-dataset/swarovski.csv')
head(auction)
head(swarovski)

dim(auction)
dim(swarovski)
summary(auction)
summary(swarovski)

levels(as.factor(auction$item))
#"Cartier wristwatch"  "Palm Pilot M515 PDA" "Xbox game console"  
levels(as.factor(auction$auction_type))
#"3 day auction" "5 day auction" "7 day auction"
```

## data manipulation

```{r}
library(tidyverse)
library(zoo)
#change characteristic varibale to dummy variables
auction$is_3day<- ifelse(auction$auction_type=='3 day auction',1,0)
auction$is_5day<- ifelse(auction$auction_type=='5 day auction',1,0)
auction$is_7day<- ifelse(auction$auction_type=='7 day auction',1,0)
auction$is_catier<- ifelse(auction$item=='Cartier wristwatch',1,0)
auction$is_palm<- ifelse(auction$item=='Palm Pilot M515 PDA',1,0)
auction$is_xbox<- ifelse(auction$item=='Xbox game console',1,0)
#add a few statistical variables
## winning bid, number of bids & average bid per auction
u<-unique(auction$auctionid)
winner<-rep(0,nrow(auction))
livebid<- rep(0,nrow(auction))
average<-rep(0,nrow(auction))
sd<- rep(0,nrow(auction))
num.bids<-rep(0,nrow(auction))

for (i in u){
        indexes<-which(auction$auctionid == i)
        cur.average<-sum(auction$bid[indexes])/length(indexes)
        cur.dev<- sd(auction$bid[indexes])
        current<- recover.livebids(auction$bid[indexes],auction$bidder[indexes],auction$openbid[indexes],auction$price[indexes])
        j<-0
        for (k in indexes){
                j<- j+1
                winner[k]<-ifelse(auction$bid[k]==auction$price[k], 1 , 0)
                num.bids[k]<-length(indexes)
                average[k]<-cur.average
                sd[k]<- cur.dev
                livebid[k]<- current[j]
        }

}
auction$winner<-winner
auction$avg.bid<-average
auction$sd.bid<- sd
auction$num.bids<-num.bids
auction$livebid<- livebid

## length of auction (number of days)
auction$length<-rep(0,nrow(auction))
idx3<-which(auction$auction_type == "3 day auction")
auction$length[idx3]<-3
idx5<-which(auction$auction_type == "5 day auction")
auction$length[idx5]<-5
idx7<-which(auction$auction_type == "7 day auction")
auction$length[idx7]<-7

## normalized bidding time (when bid was placed over auction length)
auction$time.fraction<-auction$bidtime/auction$length

## normalized bidding price (bid over winning price)
auction$bid.fraction<-auction$bid/auction$price

## bid distance from open bid
auction<-mutate(auction,dist.open=(bid-openbid)/bid)

## bid distance from average bid
auction<-mutate(auction,dist.avg=(bid-avg.bid)/avg.bid)

## add sequence of each bid
auction$num_current_bids <- ave(auction$bidtime, auction$auctionid, FUN = seq_along)

## add time difference
auction$time.diff <- ave(auction$bidtime, auction$auctionid, FUN = function(x) c(0,diff(x)))
auction$time.diff[is.na(auction$time.diff)]<- 0 

## add bid difference
auction$bid.diff <- ave(auction$bid, auction$auctionid, FUN = function(x) c(0,diff(x)))
auction$bid.diff[is.na(auction$bid.diff)]<- 0 

## add bid velocity
auction$velocity<- auction$bid.diff/auction$time.diff
auction$velocity[is.infinite(auction$velocity)]<- 0 
auction$velocity[is.nan(auction$velocity)]<- 0 

##add lag
auction$livebid.lag<- ave(auction$livebid, factor(auction$auctionid), FUN=function(x) c(NA,x[1:(length(x)-1)]))
auction$livebid.lag<- na.fill(auction$livebid.lag,0)
```

## Exploring the data

There are 627 item in the data set in total. Since the bidders do not know what others have bid and might be placing mulptiple bids after being notified that they were outbid, the last bid is not necessaily the one who get the item. 

```{r}
library(dplyr)
#get rid of the outlier
#1639672910
auction<-auction[auction$auctionid!="1639672910",]

closing_price<- auction %>% select(auctionid,openbid,price,auction_type,item)%>%
  group_by(auction$auctionid) %>% 
  summarise(openbid=max(openbid),price=max(price),auction_type=max(auction_type),item=max(item))
# text theme for every plot in this project
text_theme <- theme(text = element_text(size = 10, 
                            family = "Verdana", 
                            face = "plain"),
                    plot.title = element_text(hjust = 0.5))

# distribution of prices
ggplot(closing_price, 
       aes(x = price)) + 
  geom_histogram(binwidth = 85, 
                 aes(fill = item), 
                 col = "white") +
  geom_vline(aes(xintercept = mean(auction$price)), 
             col = "orange",
             size = 0.5) +
  labs(title = "Distribution of prices",
       x = "Final prices",
       y = "Frequency") +
  text_theme


ggplot(auction, 
       aes(x = time.fraction)) + 
  geom_histogram(binwidth = 0.01,
                 col = "white") +
  labs(title = "Distribution of Bid Time",
       x = "Bid time ",
       y = "Frequency") +
  text_theme
```

```{r}
# distribution of auction types
ggplot(closing_price, 
       aes(x = factor(1), 
           fill = closing_price$auction_type)) +
  geom_bar(width = 3) +
  ggtitle("Auctions' bid type distribution") +
  coord_polar(theta = "y") +
  labs(fill= "Type of auction") +
  scale_fill_brewer(palette = "YlGnBu") +
  xlab(NULL) + ylab(NULL) +
  text_theme
# few ggplot complements 
bar_auction_type <- geom_bar(aes(fill = auction$auction_type), 
                             stat = "count")

ggplot(closing_price, 
       aes(x = factor(1), 
           fill = item)) +
  geom_bar(width = 3) +
  ggtitle("Auctions' item distribution") +
  coord_polar(theta = "y") +
  labs(fill= "Type of auction item") +
  scale_fill_brewer(palette = "YlGnBu") +
  xlab(NULL) + ylab(NULL) +
  text_theme
# few ggplot complements 
bar_auction_type <- geom_bar(aes(fill = auction$auction_type), 
                             stat = "count")

# relationship between final price per auction and auction type
ggplot(closing_price, 
       aes(x = auction_type, y = price)) + 
  geom_boxplot(aes(fill = auction_type), 
           color = "darkgrey")  +
  labs(title = "Final price per action", 
       x = "Type of an auction", 
       y = "Final price") +
  guides(fill = F) +
  scale_fill_brewer(palette = "YlGnBu") + 
  text_theme + 
  coord_cartesian(ylim = c(0, 1000)) # few outliers masked
```

## data manipulation

```{r}
library(tidyverse)
library(zoo)
#change characteristic varibale to dummy variables
auction$is_3day<- ifelse(auction$auction_type=='3 day auction',1,0)
auction$is_5day<- ifelse(auction$auction_type=='5 day auction',1,0)
auction$is_7day<- ifelse(auction$auction_type=='7 day auction',1,0)
auction$is_catier<- ifelse(auction$item=='Cartier wristwatch',1,0)
auction$is_palm<- ifelse(auction$item=='Palm Pilot M515 PDA',1,0)
auction$is_xbox<- ifelse(auction$item=='Xbox game console',1,0)
#add a few statistical variables
## winning bid, number of bids & average bid per auction
u<-unique(auction$auctionid)
winner<-rep(0,nrow(auction))
livebid<- rep(0,nrow(auction))
average<-rep(0,nrow(auction))
sd<- rep(0,nrow(auction))
num.bids<-rep(0,nrow(auction))

for (i in u){
        indexes<-which(auction$auctionid == i)
        cur.average<-sum(auction$bid[indexes])/length(indexes)
        cur.dev<- sd(auction$bid[indexes])
        current<- recover.livebids(auction$bid[indexes],auction$bidder[indexes],auction$openbid[indexes],auction$price[indexes])
        j<-0
        for (k in indexes){
                j<- j+1
                winner[k]<-ifelse(auction$bid[k]==auction$price[k], 1 , 0)
                num.bids[k]<-length(indexes)
                average[k]<-cur.average
                sd[k]<- cur.dev
                livebid[k]<- current[j]
        }

}
auction$winner<-winner
auction$avg.bid<-average
auction$sd.bid<- sd
auction$num.bids<-num.bids
auction$livebid<- livebid

## length of auction (number of days)
auction$length<-rep(0,nrow(auction))
idx3<-which(auction$auction_type == "3 day auction")
auction$length[idx3]<-3
idx5<-which(auction$auction_type == "5 day auction")
auction$length[idx5]<-5
idx7<-which(auction$auction_type == "7 day auction")
auction$length[idx7]<-7

## normalized bidding time (when bid was placed over auction length)
auction$time.fraction<-auction$bidtime/auction$length

## normalized bidding price (bid over winning price)
auction$bid.fraction<-auction$bid/auction$price

## bid distance from open bid
auction<-mutate(auction,dist.open=(bid-openbid)/bid)

## bid distance from average bid
auction<-mutate(auction,dist.avg=(bid-avg.bid)/avg.bid)

## add sequence of each bid
auction$num_current_bids <- ave(auction$bidtime, auction$auctionid, FUN = seq_along)

## add time difference
auction$time.diff <- ave(auction$bidtime, auction$auctionid, FUN = function(x) c(0,diff(x)))
auction$time.diff[is.na(auction$time.diff)]<- 0 

## add bid difference
auction$bid.diff <- ave(auction$bid, auction$auctionid, FUN = function(x) c(0,diff(x)))
auction$bid.diff[is.na(auction$bid.diff)]<- 0 

## add bid velocity
auction$velocity<- auction$bid.diff/auction$time.diff
auction$velocity[is.infinite(auction$velocity)]<- 0 
auction$velocity[is.nan(auction$velocity)]<- 0 

##add lag
auction$livebid.lag<- ave(auction$livebid, factor(auction$auctionid), FUN=function(x) c(NA,x[1:(length(x)-1)]))
auction$livebid.lag<- na.fill(auction$livebid.lag,0)
```


```{r}
head(auction,50)
```
```{r}
auction$winner<-as.factor(auction$winner)
######## PLOTS
# text theme for every plot in this project
text_theme <- theme(text = element_text(size = 10, 
                            family = "Verdana", 
                            face = "plain"),
                    plot.title = element_text(hjust = 0.5))

# distribution of bids over item and aution type
ggplot(auction,aes(x =item, fill= winner))+
        geom_bar()+
        facet_wrap(~auction_type)+
        xlab("Auction Item")+ ylab("Total Count")+
        text_theme

# check if no. of bids has predictive power
# hypothesis - As number of bids per item increases chance of winning drops

ggplot(auction,aes(x = num.bids, fill= winner))+
        geom_histogram(binwidth = 20)+
        facet_wrap(~auction_type)+
        xlab("Number of Bids per Auction")+ ylab("Total Count")+
        text_theme

# check if distance from open bid has predictive power
# hypothesis - Bids much higher than openning bid have higher chance of winning
ggplot(auction)+
        geom_point(aes(x = dist.open,y=price, color= winner))+
        facet_wrap(~auction_type)+
        xlab("Distance form Openning Bid")+ ylab("Final Sale Price")+
        text_theme


# check if distance from average bid has predictive power
# hypothesis - Bids below the average bid have zero chance of winning

ggplot(auction)+
        geom_point(aes(x =dist.avg,y=price, color= winner))+
        facet_wrap(~auction_type)+
        xlab("Distance form Average Bid")+ ylab("Final Sale Price")+
        text_theme


# check if time fraction has predictive power
# hypothesis - Bids placed toward the end of the auction have higher chance of winning

ggplot(auction)+
        geom_point(aes(x =time.fraction,y=price, color= winner))+
        facet_wrap(~auction_type)+
        xlab("Bidding Time")+ ylab("Final Sale Price")+
        text_theme

```

### modeling a single auction using Smoothing splines
```{r}
#Example of a 7-day auction
sevenday<- auction[which(auction$length==7),]
ggplot(sevenday, aes(x=bidtime, y=bid, color= item)) +
   geom_point() + 
   scale_fill_brewer(palette = "YlGnBu") + 
   xlab("bid time") +
   ggtitle("Scatter Plot of 7-day auctions")+
   expand_limits(x=7)+
   text_theme

sample3<- auction[which(auction$auctionid=='1639226378'),]
ggplot(sample3, aes(x=bidtime, y=bid)) +
   geom_line() + 
   xlab("bid time") +
   ggtitle("Sample Auction 1639226378")+
   expand_limits(x=7)+
   text_theme

ggplot(sample3, aes(x=bidtime, y=livebid)) +
   geom_line() + 
   xlab("bid time") +
   ggtitle("Sample Auction 1639226378")+
   expand_limits(x=7)+
   text_theme
```

## model one singel auction using non-linear polynomial regression
```{r}
library(splines)
library(gam)

rss <- rep(0, 7)
for (i in 1:7) {
    fit <- lm(bid ~ poly(bidtime, i), data = sample3)
    rss[i] <- sum(fit$residuals^2)
}
plot(1:7, rss, xlab = "Degree", ylab = "RSS", type = "l", col='red')


fit.lm <- lm(bid ~ poly(bidtime,3), data = sample3)
summary(fit.lm)

bidtime.new <- seq(1, 7, by = 0.01)
pred <- predict(fit.lm, list(bidtime = bidtime.new))
plot(bid ~ bidtime, data = sample3, col = "darkgrey")
lines(bidtime.new,pred, col = "red", lwd = 2)
```
The RSS decreases as the degree of the polynomial decreases. It achieves its minimum at degree 7. However, it diverges as the degree of polynomial increases. Thus from the plot we choose the degree of 3. 

## Soomthing splines and gam
```{r}
rss <- rep(0, 16)
for (i in 3:16) {
    fit <- lm(bid ~ bs(bidtime, df = i), data = sample3)
    rss[i] <- sum(fit$residuals^2)
}
plot(3:16, rss[-c(1, 2)], xlab = "Degrees of freedom", ylab = "RSS", type = "l")


fit.spline <- lm( bid~ bs(bidtime, knots = c(5, 14)), data = sample3)
summary(fit.spline)

pred.spline <- predict(fit.spline, list(bidtime = bidtime.new))
plot(bid ~ bidtime, data = sample3, col = "darkgrey")
lines(bidtime.new, pred.spline, col = "orchid", lwd = 2)
```
```{r}
bid<- sample3$bid
bidtime<- sample3$bidtime
fit.sspline <- smooth.spline(bid, bidtime,df=3)
fit.gam<- gam(bid ~ s(bidtime,7))


pred.sspline<- predict(fit.sspline, bidtime.new)
pred.gam <- predict(fit.gam, list(bidtime = bidtime.new))
plot(bid ~ bidtime, data = sample3, col = "darkgrey")
lines(bidtime.new,pred.sspline$y, col = "red", lwd = 2)
lines(bidtime.new,pred.gam, col = "blue", lwd = 2)

summary(fit.gam)
```
## Add bidding pattern

```{r}
#split by auction
duration<-1
sLive <- split(auction$livebid,auction$auctionid)
sTime <- split(auction$time.fraction,auction$auctionid)
#create variables with opening and closing bid added
numauct<-length(sLive)
sLiveEnds<-list(0);length(sLiveEnds)<-numauct
sTimeEnds<-list(0); length(sTimeEnds)<-numauct
sLiveStart<-list(0); length(sLiveStart)<-numauct
sTimeStart<-list(0); length(sTimeStart)<-numauct
sTimeEndsTemp<-list(0); length(sTimeStart)<-numauct
sTimeStartTemp<-list(0); length(sTimeStartTemp)<-numauct
for (j in 1:numauct){
     sLiveEnds[[j]]<-c(min(sLive[[j]]),sLive[[j]],max(sLive[[j]]))
     sTimeEnds[[j]]<-c(0,sTime[[j]],duration)
     sLiveStart[[j]]<-c(min(sLive[[j]]),sLive[[j]])
     sTimeStart[[j]]<-c(0,sTime[[j]])
     sTimeEndsTemp[[j]]<-c(0.000001,sTime[[j]],duration)
     sTimeStartTemp[[j]]<-c(0.000001,sTime[[j]])
     }

delta<- 0.01
epsilon<-0.000001
num.mod<-4  ##number of models
w.x<-.5
w.y<-.5


########################
##SSE for Diffent Models
########################
n<-length(sLive)  ##number of auctions I'm looking at 
wSSE.range<-array(0,c(n,num.mod))  
wSSE.var<-array(0,c(n,num.mod))  
models.range<-array(0,n)
models.var<-array(0,n)

for (i in 1:n){

## Exponential Model (Model 1)
x<-seq(0,duration,by=.01)
reg1<-lm(log(sLiveEnds[[i]])~ sTimeEnds[[i]])
fit1y<-exp(as.vector(reg1$coefficients)[1])*exp(as.vector(reg1$coefficients)[2]*sTimeEnds[[i]])
fit1x<-log(sLiveEnds[[i]]/exp(as.vector(reg1$coefficients)[1]))/as.vector(reg1$coefficients)[2]
fit1.SSEy<- sum((fit1y-sLiveEnds[[i]])^2)
fit1.SSEx<- sum((fit1x-sTimeEnds[[i]])^2)
wSSE.range[i,1]<- (w.y*fit1.SSEy)/diff(range(sLive[[i]]))^2 + (w.x*fit1.SSEx)/diff(range(sTimeEnds[[i]]))^2
wSSE.var[i,1]<- (w.y*fit1.SSEy)/var(sLive[[i]]) + (w.x*fit1.SSEx)/var(sTimeEnds[[i]])

## Invere Exponential Model (Model 2) 
x<-seq(0,max(sLive[[i]]),by=.1)
reg2<-lm(log(sTimeEndsTemp[[i]])~ sLiveEnds[[i]])
fit2y<-log(sTimeEndsTemp[[i]]/exp(as.vector(reg2$coefficients)[1]))  /as.vector(reg2$coefficients)[2]
fit2x<-exp(as.vector(reg2$coefficients)[1]) * exp(as.vector(reg2$coefficients)[2]*sLiveEnds[[i]])
fit2.SSEy<- sum((fit2y-sLiveEnds[[i]])^2)
fit2.SSEx<- sum((fit2x-sTimeEnds[[i]])^2)
wSSE.range[i,2]<- (w.y*fit2.SSEy)/diff(range(sLive[[i]]))^2 + (w.x*fit2.SSEx)/diff(range(sTimeEnds[[i]]))^2
wSSE.var[i,2]<- (w.y*fit2.SSEy)/var(sLive[[i]]) + (w.x*fit2.SSEx)/var(sTimeEnds[[i]])

## Logistic Growth (Model 3) 
x<-seq(0,duration,by=.01)
reg3<- lm( log((max(sLiveStart[[i]])+delta)/sLiveStart[[i]]-1) ~ sTimeStart[[i]] )
fit3y<- (max(sLiveEnds[[i]])+delta)/(1+exp(as.vector(reg3$coefficients)[1])*exp(as.vector(reg3$coefficients)[2]*sTimeEnds[[i]]))
fit3x<- (log( (max(sLiveEnds[[i]])+delta)/sLiveEnds[[i]] - 1) - log (exp(as.vector(reg3$coefficients)[1])))/as.vector(reg3$coefficients)[2]
fit3.SSEy<- sum((fit3y-sLiveEnds[[i]])^2)
fit3.SSEx<- sum((fit3x-sTimeEnds[[i]])^2)
wSSE.range[i,3]<- (w.y*fit3.SSEy)/diff(range(sLive[[i]]))^2 + (w.x*fit3.SSEx)/diff(range(sTimeEnds[[i]]))^2
wSSE.var[i,3]<- (w.y*fit3.SSEy)/var(sLive[[i]]) + (w.x*fit3.SSEx)/var(sTimeEnds[[i]])

## Inverse Logistic Growth (Model 4)
x<-seq(0,max(sLive[[i]]),by=.1)
reg4<- lm( log((duration+epsilon)/sTimeEndsTemp[[i]]-1) ~ sLiveEnds[[i]])
fit4y<- (log(((duration+epsilon)/sTimeEndsTemp[[i]])-1) - as.vector(reg4$coefficients)[1]) /as.vector(reg4$coefficients)[2]
fit4x<- (duration+epsilon)/(1+exp(as.vector(reg4$coefficients)[1])*exp(as.vector(reg4$coefficients)[2]*sLiveEnds[[i]]))
fit4.SSEy<- sum((fit4y-sLiveEnds[[i]])^2)
fit4.SSEx<- sum((fit4x-sTimeEnds[[i]])^2)
wSSE.range[i,4]<- (w.y*fit4.SSEy)/diff(range(sLive[[i]]))^2 + (w.x*fit4.SSEx)/diff(range(sTimeEnds[[i]]))^2
wSSE.var[i,4]<- (w.y*fit4.SSEy)/var(sLive[[i]]) + (w.x*fit4.SSEx)/var(sTimeEnds[[i]])

}

```

```{r error=TRUE}
for (k in 1:n){
     wSSE.range[k,][wSSE.range[k,]=="NaN"]=(min(na.omit(wSSE.range[k,])+1))
     wSSE.var[k,][wSSE.var[k,]=="NaN"]=(min(na.omit(wSSE.var[k,])+1))
     models.range[k]<-(1:num.mod)[min(wSSE.range[k,])==wSSE.range[k,]]
     models.var[k]<-(1:num.mod)[min(wSSE.var[k,])==wSSE.var[k,]]
}
(1:n)[models.range!=models.var]
table(models.range)
table(models.var)

numauct<-length(sLive)
x<-seq(0,duration,by=.1)
x[1]<-min(x)+.001
x[length(x)]<-max(x)-.01
numplot<-length(x)
ypred <- yfdpred7 <- ysdpred7 <- array(0,c(numauct,numplot))

models<- models.range


################################################################
##Creating Fitted Data Based on Selected (by SSE criteria) Model
################################################################
## Matrix of data
numauct<-length(sLive)
x<-seq(0,duration,by=.05)
x[1]<-min(x)+.001
x[length(x)]<-max(x)-.01
numplot<-length(x)
ypred <- yfdpred7 <- ysdpred7 <- array(0,c(numauct,numplot))

models7<- models<- models.range   ###IMPT: THIS IS THE SSE CRITERIA THAT I AM USING (either models.range or models.var) 

for (i in 1:n){
if (models[i]==1){
reg1<-lm(log(sLiveEnds[[i]])~ sTimeEnds[[i]])
ypred[i,]<-exp(as.vector(reg1$coefficients)[1])*exp(as.vector(reg1$coefficients)[2]*x)
yfdpred7[i,]<-exp(as.vector(reg1$coefficients)[1])*as.vector(reg1$coefficients)[2]*exp(as.vector(reg1$coefficients)[2]*x)
ysdpred7[i,]<-exp(as.vector(reg1$coefficients)[1])*(as.vector(reg1$coefficients)[2])^2*exp(as.vector(reg1$coefficients)[2]*x)
}

if (models[i]==2){
reg2<-lm(log(sTimeEndsTemp[[i]])~ sLiveEnds[[i]])
ypred[i,]<-log(x/exp(as.vector(reg2$coefficients)[1]))  /as.vector(reg2$coefficients)[2]
yfdpred7[i,]<-1/(as.vector(reg2$coefficients)[2]*x)
ysdpred7[i,]<- -1/(as.vector(reg2$coefficients)[2]*x^2)
}

if (models[i]==3){
reg3<- lm( log((max(sLiveStart[[i]])+delta)/sLiveStart[[i]]-1) ~ sTimeStart[[i]] )
ypred[i,]<- (max(sLive[[i]])+delta)/(1+exp(as.vector(reg3$coefficients)[1])*exp(as.vector(reg3$coefficients)[2]*x))
yfdpred7[i,]<- -(max(sLive[[i]])+delta)*exp(as.vector(reg3$coefficients)[1])*as.vector(reg3$coefficients)[2]*exp(as.vector(reg3$coefficients)[2]*x)/
	  (1+exp(as.vector(reg3$coefficients)[1])*exp(as.vector(reg3$coefficients)[2]*x))^2
ysdpred7[i,]<- -(max(sLive[[i]])+delta)*exp(as.vector(reg3$coefficients)[1])*(as.vector(reg3$coefficients)[2])^2*exp(as.vector(reg3$coefficients)[2]*x)*
           (1-exp(as.vector(reg3$coefficients)[1])*exp(as.vector(reg3$coefficients)[2]*x))/
	   (1+exp(as.vector(reg3$coefficients)[1])*exp(as.vector(reg3$coefficients)[2]*x))^3
}

if (models[i]==4){
reg4<- lm( log((duration+epsilon)/sTimeEndsTemp[[i]]-1) ~ sLiveEnds[[i]])
ypred[i,]<- (log(((duration+epsilon)/x)-1) - as.vector(reg4$coefficients)[1]) /as.vector(reg4$coefficients)[2]
yfdpred7[i,]<- -(duration+epsilon)/(as.vector(reg4$coefficients)[2]*(x^2)*(((duration+epsilon)/x)-1))
ysdpred7[i,]<- ((duration+epsilon)*((duration+epsilon)-2*x))/(as.vector(reg4$coefficients)[2]*(x^4)*(((duration+epsilon)/x)-1)^2)
}
}
```

```{r}
plot(x,ypred[1,],type="l",xlim=c(0,1),ylim=c(0,max(ypred)),xlab="Bid Time",ylab="Price",col='red')
for (i in 2:numauct)
lines(x,ypred[i,], lty=3)

lines(x,ypred[86,], col='yellow')
lines(x,ypred[61,], col='green')
lines(x,ypred[9,], col='purple')

legend(0, 2800, legend=c("model 1", "model 2", "model 3", "model 4"),
       col=c("red", "yellow", 'green', 'purple'), lty=1, cex=0.8)

table(models)
```


## Static model

```{r}
library(dplyr)
library(caret)
# Split data into training and testing
set.seed(123)
Train <- createDataPartition(u, p = 0.7, list = FALSE)
Train.auction<- auction$auctionid %in% u[Train]
training <- auction[Train.auction,]
testing <- auction[!Train.auction,]
```
## static prediction

Static competition features are known at the outset of the auction and do not change during the auction;
- opening price of concurrent auctions
- duration of concurrent auctions
- sellers’ feedback of concurrent auctions
```{r}
library(ggplot2)
library(caret)

#######Static Model Predicting the End Price######
set.seed(123)
train_control<- trainControl(method = 'cv',number=10,repeats = 3)

#fit auction with linear model
static.lm <- train(price~ openbid+length+bidderrate+is_catier+is_palm+is_xbox+is_3day+is_5day+is_7day, data=training, trControl=train_control, method="lm",na.action=na.exclude)
static.lm

static.pred.lm<- predict(static.lm, testing)
postResample(pred = static.pred.lm, obs = testing$livebid)

##fit auction with knn model
static.knn<- train(price ~ openbid+length+bidderrate+is_catier+is_palm+is_xbox+is_3day+is_5day+is_7day, data=training, trControl=train_control, method="knn",na.action=na.exclude)
static.knn
plot(static.knn)

static.pred.knn<- predict(static.knn, testing)
postResample(pred = static.pred.knn, obs = testing$livebid)

```

```{r}
library(ggplot2)
library(caret)

#######Static Model######
set.seed(123)

train_control<- trainControl(method = 'cv',number=10,repeats = 3)

#fit auction with linear model
static.lm <- train(livebid ~ openbid+length+bidderrate+is_catier+is_palm+is_xbox+is_3day+is_5day+is_7day, data=training, trControl=train_control, method="lm",na.action=na.exclude)
static.lm

static.pred.lm<- predict(static.lm, testing)
postResample(pred = static.pred.lm, obs = testing$livebid)

##fit auction with knn model
static.knn<- train(livebid ~ openbid+length+bidderrate+is_catier+is_palm+is_xbox+is_3day+is_5day+is_7day, data=training, trControl=train_control, method="knn",na.action=na.exclude)
static.knn
plot(static.knn)

static.pred.knn<- predict(static.knn, testing)
postResample(pred = static.pred.knn, obs = testing$livebid)

```

```{r}
#######Static + Evolving info Model######
#current price, time left, current number of bids, current number of bidders

#fit auction with linear model
se.lm <- train(livebid ~ openbid+length+bidderrate+is_catier+is_palm+is_xbox+is_3day+is_5day+is_7day+livebid.lag+num_current_bids+avg.bid+sd.bid+num.bids+time.fraction, data=training, trControl=train_control, method="lm",na.action=na.exclude)
static.lm

se.pred.lm<- predict(se.lm, testing)
postResample(pred = se.pred.lm, obs = testing$livebid)

##fit auction with knn model
se.knn<- train(livebid ~ openbid+length+bidderrate+is_catier+is_palm+is_xbox+is_3day+is_5day+is_7day+livebid.lag+num_current_bids+avg.bid+sd.bid+num.bids+time.fraction, data=training, trControl=train_control, method="knn",na.action=na.exclude)
static.knn
plot(static.knn)

se.pred.knn<- predict(se.knn, testing)
postResample(pred = se.pred.knn, obs = testing$livebid)
```


```{r}
#######Static + Evolving info + dynamic Model######
#distance to 

#fit auction with linear model
sed.lm <- train(livebid ~ openbid+length+bidderrate+is_catier+is_palm+is_xbox+is_3day+is_5day+is_7day+livebid.lag+num_current_bids+avg.bid+sd.bid+num.bids+time.fraction+bid.fraction+dist.open+dist.avg+time.diff+bid.diff+velocity, data=training, trControl=train_control, method="lm",na.action=na.exclude)
static.lm

sed.pred.lm<- predict(sed.lm, testing)
postResample(pred = sed.pred.lm, obs = testing$livebid)

##fit auction with knn model
sed.knn<- train(livebid ~ openbid+length+bidderrate+is_catier+is_palm+is_xbox+is_3day+is_5day+is_7day+livebid.lag+num_current_bids+avg.bid+sd.bid+num.bids+time.fraction+bid.fraction+dist.open+dist.avg+time.diff+bid.diff+velocity, data=training, trControl=train_control, method="knn",na.action=na.exclude)
static.knn
plot(static.knn)

sed.pred.knn<- predict(sed.knn, testing)
postResample(pred = sed.pred.knn, obs = testing$livebid)
```


```{r error=TRUE}
library(ggplot2)
library(xgboost)
library(caret)

#######XGboost Full model######
set.seed(123)

train_x<- as.matrix(training %>% select(-which(sapply(.,is.character))) %>%select(-c(livebid,auctionid)))

train_y<- training$livebid

test_x<- as.matrix(testing %>% select(-which(sapply(.,is.character))) %>%select(-c(livebid,auctionid)))

test_y<- testing$livebid

full.xg<- xgboost(data = train_x,
                    label=train_y,
                 booster = "gblinear",
                 objective = "reg:linear",
                 max.depth = 8,
                 nround = 500,
                 lambda = 0,
                 lambda_bias = 0,
                 alpha = 0, 
                 missing=NA, 
                 verbose = 0)


pred.test.xgboost <- predict(full.xg,test_x, missing=NA)
rmse <- function(error)
{
    sqrt(mean(error^2))
}
 
# Function that returns Mean Absolute Error
mae <- function(error)
{
    mean(abs(error))
}

error<- test_y-pred.test.xgboost

postResample(pred = pred.test.xgboost, obs = testing$livebid)
```

## using LSTM to predicting the bid pattern
```{r error=TRUE}
library(keras)

lags <- function(x, k){
    
    lagged =  c(rep(NA, k), x[1:(length(x)-k)])
    DF = as.data.frame(cbind(lagged, x))
    colnames(DF) <- c( paste0('x-', k), 'x')
    DF[is.na(DF)] <- 0
    return(DF)
  }

lstm.train = lags(training$livebid, 1)
lstm.test = lags(testing$livebid, 1)


normalize <- function(train, test, feature_range = c(0, 1)) {
    x = train
    fr_min = feature_range[1]
    fr_max = feature_range[2]
    std_train = ((x - min(x) ) / (max(x) - min(x)  ))
    std_test  = ((test - min(x) ) / (max(x) - min(x)  ))
    
    scaled_train = std_train *(fr_max -fr_min) + fr_min
    scaled_test = std_test *(fr_max -fr_min) + fr_min
    
    return( list(scaled_train = as.vector(scaled_train), scaled_test = as.vector(scaled_test) ,scaler= c(min =min(x), max = max(x))) )
    
  }

  
Scaled = normalize(train, test, c(-1, 1))
  
y_train = lstm.train[, 2]
x_train = lstm.train[, 1]
y_test = lstm.test[, 2]
x_test = lstm.test[, 1]
  
dim(x_train) <- c(length(x_train), 1, 1)
dim(x_train)
batch_size = 1
units = 1
# specify required arguments
model <- keras_model_sequential() %>% 
  layer_lstm(units=128,batch_input_shape = c(batch_size, dim(x_train)[2], dim(x_train)[3]),   stateful= TRUE)%>%
  layer_dense(units = 32) %>% 
  layer_dropout(0.2) %>% 
  layer_dense(units = 1)

model %>% compile(
  optimizer_adam( lr= 0.02 , decay = 1e-6 ),
  loss = "mse",
  metrics = list("mean_absolute_error")
)



summary(model)


history <- model %>% fit(
  x_train,y_train,
  epochs = 10,
  batch_size = 1
)

```

```{r error=TRUE}
dim(x_test) <- c(length(x_test), 1, 1)
yhat = model %>% predict(X, batch_size=1)

postResample(pred = yhat, obs = y_test)
```

evolving competition features change during the auction, such as the 
current price of concurrent auctions (if the price is low in other auctions, bidders may leave the focal auction) or 
the number of bidders in concurrent auctions (bidders may feel that their chances of winning are higher in auctions with lower competition)


price dynamic competition features capture the effect of changing dynamics in competing auctions



