---
title: "Math642_HW11_FyonaSun"
author: "Fyona Sun"
date: "4/22/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1.
```{r}
source("DBDA2E-utilities.R")  # Load definitions of graphics functions etc.
source("BernBeta.R") 

# Specify the prior:
t = 0.75             # Specify the prior MODE.
n = 25               # Specify the effective prior sample size.
a = t*(n-2) + 1      # Convert to beta shape parameter a.
b = (1-t)*(n-2) + 1  # Convert to beta shape parameter b.

Prior = c(a,b)       # Specify Prior as vector with the two shape parameters.

# Specify the data:
N = 25                         # The total number of flips.
z = 1                         # The number of heads.
Data = c(rep(0,N-z),rep(1,z))  # Convert N and z into vector of 0's and 1's.
# 
# openGraph(width=5,height=7)
# posterior = BernBeta( priorBetaAB=Prior, Data=Data , plotType="Bars" , 
#                       showCentTend="Mode" , showHDI=TRUE , showpD=FALSE )
# saveGraph(file="BernBetaExample",type="png")
```
a) What are the inputs and outputs for BernBeta?  Explain what they mean
.
Input of BernBeta is prior of a beta distribution and new observations.
priorBetaAB is two-element vector of beta distribution B(a,b).
Data is vector of 0's and 1's. z is the number of observed heads from the data and N is the total number of trials from the data.

The output is a two-element vector that gives posterior beta parameters
  
b) Start with a prior of beta(4,4).  Use the BernBeta function to flip the coin once and get a head.  What is the posterior?  Show all plots and explain results.

```{r}
BernBeta(priorBetaAB = c(4, 4), Data = c(1), plotType="Bars", showCentTend="Mode" , showHDI=TRUE , showpD=FALSE )
```
The posterior is B(5,4). The observed head shifts the prior disstribution slightly to the right. The Posterior is a compromise between our prior beliefs and a single data point observed.

c) Use the posterior from the previous flip as the prior for the next flip.  (You can use the result from the last run - “post” - instead of c(5,4) as the input the BernBeta.) Flip the coin again and get a head again.  What is the new posterior?
```{r}
BernBeta(priorBetaAB = c(5, 4), Data = c(1), plotType="Bars", showCentTend="Mode" , showHDI=TRUE , showpD=FALSE )
```

The new posterior is B(6,4).

d) Using the posterior from the last flip, flip again and get tails.  What is the new posterior?

```{r}
BernBeta(priorBetaAB = c(6, 4), Data = c(0), plotType="Bars", showCentTend="Mode" , showHDI=TRUE , showpD=FALSE )
```
The new posterior is B(6,5).


e) Start over at b) and do three flips, but in the order T, H, H.  Show the posterior after these 3 flips. How does the posterior change?  Does order of the results matter?
```{r}
BernBeta(priorBetaAB = c(4, 4), Data = c(0,1,1), plotType="Bars", showCentTend="Mode" , showHDI=TRUE , showpD=FALSE )

BernBeta(priorBetaAB = c(4, 4), Data = c(1,0,1), plotType="Bars", showCentTend="Mode" , showHDI=TRUE , showpD=FALSE )

BernBeta(priorBetaAB = c(4, 4), Data = c(1,1,0), plotType="Bars", showCentTend="Mode" , showHDI=TRUE , showpD=FALSE )
```
We get the same final posterior form as before regardless of ordering of the 3 coin tosses. The posterior distribution is B(6,5)

f) Start over at b) and do three flips, but in the order T, T, H.  Show the posterior after these 3 flips. How does the posterior change?
```{r}
BernBeta(priorBetaAB = c(4, 4), Data = c(0,0,1), plotType="Bars", showCentTend="Mode" , showHDI=TRUE , showpD=FALSE )
```
The posterior distribution is B(5,6)
